---
title: 'Project: Question 7'
author: 'Gabriel Berardi'
output: pdf_document
---

```{r chunk 1, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(knitr)

data <- load('airbnb.RData')
rm(amsterdam, copenhagen, dublin, iceland, london, paris, warsaw)

bande <- rbind(barcelona, edinburgh)
```

# 1. Two-proportion Test to Compare the Proportion of Properties with Room Type *"Private room"*

Lastly, we are going to conduct a two-proportion test to formally compare the proportion of properties with room type of *Private room* between Barcelona and Edinburgh.

First of all, we should calculate how many of the listings in Barcelona and Edinburgh are "private rooms" and "non private rooms":

```{r chunk 2, echo = FALSE}

bande_table <- bande %>% 
  mutate(room_type = factor(ifelse(room_type == 'Private room',
                                   'Private room',
                                   'Non Private room')))


bande_table <- table(bande_table$city, bande_table$room_type)
bande_table
```

Let's visualize this:

```{r chunk 3, echo = FALSE, fig.height = 4}

ggplot(as.data.frame(bande_table), aes(x=Var1, y = Freq, fill=Var2)) + 
    geom_bar(stat='identity') +
  ggtitle('Proportin of Room Type in Barcelona and Edinburgh') +
  xlab('') +
  ylab('Count') +
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = 'bold'),
        axis.title = element_text(size = 8, face = 'bold'),
        legend.position = 'bottom',
        legend.title = element_blank())
```

\newpage

Then, we can conduct the two-proportion test:

```{r chunk 4, echo = FALSE}

prop.test(bande_table)
```

This provides a 95 % confidence interval for the difference between the proportion of the private rooms in Barcelona and Edinburgh. Since the p-value is well below 5 % and the confidence interval does not include 0 then there is sufficient evidence of a difference in the proportion of private rooms in Barcelona and Edinburgh. This difference is likely to lie between 0.12 and 0.09, with Barcelona having a higher proportion of private rooms compared to Edinburgh.
